
```{r}
library(dplyr)
library(scales)
library(tidyr)
library(reshape2)
library(ggplot2)
library(factoextra)
library(caret)
library(readr)
library(e1071)
library(rpart)
library(Metrics)
library(randomForest)
```

```{r}
data <- read.csv(file = "cleaned_dataset1.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
```


```{r}
View(data)
```



```{r}
set.seed(123)  # Set a seed for reproducibility
sampled_rows <- sample(nrow(data), 5000, replace = FALSE)
df_subset <- data[sampled_rows, ]

lapply(df_subset[sapply(df_subset,is.numeric)], shapiro.test)
```

# Univariate Analysis


```{r}
ggplot(data, aes(x=avg_price_per_room)) + 
  geom_histogram(binwidth = 9, color= 'black', fill='darkgreen',alpha = 0.8) + 
  ggtitle("Histogram of Average Room Price") +
  xlab("average_price_per_room") + 
  ylab("Number of Bookings (sqrt Scaled)") +
  scale_y_sqrt(breaks = breaks_extended(15)) +
  scale_x_continuous(breaks=seq(min(data$avg_price_per_room), 
                                max(data$avg_price_per_room), 
                                by=50)) +
  theme(plot.title = element_text(hjust = 0.5))
```



```{r}
table(as.factor(data$no_of_adults))
```


```{r}
ggplot(data, aes(x=no_of_adults)) + geom_bar(color = 'black',fill='darkgreen',alpha = 0.9) +
  xlab("no_of_adults") + 
  ylab("Number of Bookings (sqrt Scaled)") +
  scale_y_sqrt(breaks = breaks_extended(10)) +
  ggtitle("Histogram of Number of Adults") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
table(as.factor(data$no_of_children))
```


```{r}
ggplot(data, aes(x=no_of_children)) + 
  geom_bar(color = 'black',fill='darkgreen',alpha = 0.9) + 
  xlab("no_of_children") + 
  ylab("Number of Bookings (sqrt Scaled)") +
  scale_y_sqrt(breaks = breaks_extended(10)) +
  scale_x_continuous(breaks=seq(min(data$no_of_children), 
                                max(data$no_of_children), 
                                by=1)) +
  ggtitle("Histogram of Number of Children") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
table(as.factor(data$no_of_weekend_nights))
```


```{r}
ggplot(data, aes(x=factor(no_of_weekend_nights))) + 
  geom_bar(color = 'black',fill='darkgreen',alpha = 0.9) + 
  xlab("no_of_weekend_nights") + 
  ylab("Number of Bookings (sqrt Scaled)") + 
  scale_y_sqrt(breaks = breaks_extended(10)) +
  ggtitle("Histogram of Number of Weekend Nights") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
table(as.factor(data$no_of_week_nights))
```

```{r}
ggplot(data, aes(x=factor(no_of_week_nights))) + 
  geom_bar(color = 'black',fill='darkgreen',alpha = 0.8) + 
  xlab("no_of_week_nights") + 
  ylab("Number of Bookings (sqrt Scaled)") + 
  scale_y_sqrt(breaks = breaks_extended(15)) +
  ggtitle("Histogram of Number of Week Nights") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data, aes(x=lead_time)) + geom_histogram(binwidth = 10, color = 'black',fill='darkgreen',alpha = 0.8) +
  xlab("lead_time") + 
  ylab("Number of Bookings") +
  scale_x_continuous(breaks=seq(min(data$lead_time), 
                                max(data$lead_time), 
                                by=50)) +
  ggtitle("Histogram of Lead Time") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
table(as.factor(data$no_of_previous_cancellations))
```


```{r}
ggplot(data, aes(x=no_of_previous_cancellations)) +
  geom_bar(color = 'black', fill='darkgreen', alpha = 0.85) +
  xlab("no_of_previous_cancellations") +
  ylab("Number of Bookings (log10 scaled)") +
  scale_y_log10() + 
  scale_x_continuous(breaks=seq(min(data$no_of_previous_cancellations), 
                                max(data$no_of_previous_cancellations), 
                                by=1)) +
  ggtitle("Histogram of Number of Previous Cancellations") +
  theme(plot.title = element_text(hjust = 0.5))
```



```{r}
ggplot(data, aes(x=no_of_previous_bookings_not_canceled)) + geom_bar(width=0.9, color = 'black',fill='darkgreen',alpha = 0.8) +   xlab("no_of_previous_bookings_not_canceled") +   
  ylab("Number of Bookings (log10 scaled)") +
  scale_y_log10() +
  scale_x_continuous(breaks=seq(min(data$no_of_previous_bookings_not_canceled), 
                                max(data$no_of_previous_bookings_not_canceled), 
                                by=5)) +
  ggtitle("Histogram of Number of not Cancelled Previous Bookings") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
table(as.factor(data$no_of_special_requests))
```


```{r}
ggplot(data, aes(x=factor(no_of_special_requests ))) + 
  geom_bar(color = 'black',fill='darkgreen',alpha = 0.9) + 
  xlab("no_of_special_requests ") + 
  ylab("Number of Bookings (sqrt Scaled)") +
  scale_y_sqrt(breaks = breaks_extended(15))  +
  ggtitle("Histogram of Number of Special Requests") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data, aes(x=factor(arrival_year))) + 
  geom_bar(width = 0.7, color = 'black',fill='darkgreen',alpha = 0.9) + 
  xlab("arrival_year") + 
  ylab("Number of Bookings") +
  scale_y_continuous(breaks=c(6514, 29761))  +
  ggtitle("Histogram of Arrival Year") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data, aes(x=factor(arrival_month))) + 
  geom_bar(width=0.925,color = 'black',fill='darkgreen',alpha = 0.85) + 
  xlab("arrival_month") + 
  ylab("Number of Bookings") +
  scale_y_continuous(breaks = seq(0, 8000, by=1000)) +
  ggtitle("Histogram of Arrival Month") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data, aes(x=factor(arrival_date))) + 
  geom_bar(width=0.9,color = 'black',fill='darkgreen',alpha = 0.85) + 
  xlab("arrival_date") + 
  ylab("Number of Bookings") +
  scale_y_continuous(breaks = c(0,200,400,600,800,1000,1200,1400)) +
  ggtitle("Histogram of Arrival Date") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data, aes(x=type_of_meal_plan )) + 
  geom_bar(color = 'black',fill='darkgreen', alpha = 0.9) + 
  xlab("Type of Meal Plan") + 
  ylab("Number of Bookings (sqrt Scaled)") +
  scale_y_sqrt(breaks = breaks_extended(10)) +
  ggtitle("Histogram of Meal Plan Type") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data, aes(x=factor(required_car_parking_space))) + 
  geom_bar(color = 'black',fill='darkgreen',alpha = 0.9) + 
  xlab("Car Parking Required") +
  ylab("Number of Bookings(sqrt Scaled)") +
  scale_y_sqrt(breaks = c(35147, 1128)) +
  ggtitle("Histogram of Car Parking Requirement") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data, aes(x = room_type_reserved)) +
  geom_bar(stat = "count", color = 'black', fill = 'darkgreen', alpha = 0.85) +
  xlab("Room Type") +
  ylab("Number of Bookings (sqrt Scaled)") +
  scale_y_sqrt(breaks = breaks_extended(10)) +
  ggtitle("Histogram of Room Type") +
  theme(plot.title = element_text(hjust = 0.5))
```



```{r}
table(as.factor(data$repeated_guest))
```


```{r}
ggplot(data, aes(x=factor(repeated_guest))) + 
  geom_bar(color = 'black',fill='darkgreen',alpha = 0.9) + 
  xlab("Repeated Guest") +
  ylab("Number of Bookings (sqrt Scaled)") +
  scale_y_sqrt(breaks = c(35344, 930)) +
  ggtitle("Histogram of Repeated Guest") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
table(as.factor(data$booking_status))
```

```{r}
ggplot(data, aes(x=booking_status)) + 
  geom_bar(color = 'black',fill='darkgreen',alpha = 0.9) +
  xlab("Booking Status") +
  ylab("Number of Bookings") +
  scale_y_continuous(breaks=c(11885, 24389)) +
  ggtitle("Histogram of Booking Status") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
names(data[sapply(data, is.numeric)])
```

```{r}
ggplot(data, aes(x=no_of_adults, y=avg_price_per_room)) +
  geom_point(shape=1,size=1.5, colour="#D17A00", stroke = 0.4) +
  geom_smooth(method = "lm") + 
  xlab("no_of_adults") + 
  ylab("avg_price_per_room") +
  ggtitle("Scatter Plot of Number of Adults vs Average Room Price") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data, aes(x=no_of_children, y=avg_price_per_room)) +
  geom_point(shape=1,size=1.5, colour="#D17A00", stroke = 0.5) +
  geom_smooth(method = "lm") + 
  xlab("no_of_children") + 
  ylab("avg_price_per_room") +
  scale_x_continuous(breaks=seq(c(1,2,3,5,6,7,8,9,10,11))) +
  ggtitle("Scatter Plot of Number of Children vs Average Room Price") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
subset(data, no_of_children > 5)
```

```{r}
ggplot(data, aes(x=no_of_weekend_nights, y=avg_price_per_room)) +
  geom_point(shape=1,size=1.5, colour="#D17A00", stroke = 0.4) +
  geom_smooth(method = "lm") + 
  xlab("no_of_weekend_nights") + 
  ylab("avg_price_per_room") +
  ggtitle("Scatter Plot of Number of Weekend Nights vs Average Room Price") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data, aes(x=no_of_week_nights, y=avg_price_per_room)) +
  geom_point(shape=1,size=1.5, colour="#D17A00", stroke = 0.4) +
  geom_smooth(method = "lm") + 
  xlab("no_of_week_nights") + 
  ylab("avg_price_per_room") +
  ggtitle("Scatter Plot of Number of Week Nights vs Average Room Price") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data, aes(x=lead_time, y=avg_price_per_room)) +
  geom_point(shape=1,size=1, colour="#D17A00", stroke = 0.35) +
  geom_smooth(method = "lm") + 
  xlab("lead_time") + 
  ylab("avg_price_per_room") +
  ggtitle("Scatter Plot of Lead Time vs Average Room Price") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data, aes(x=arrival_month, y=avg_price_per_room)) +
  geom_point(shape=1,size=1.5, colour="#D17A00", stroke = 0.5) +
  geom_smooth(method = "lm") + 
  xlab("arrival_month") + 
  ylab("avg_price_per_room") +
  scale_x_continuous(breaks=seq(c(1,2,3,4,5,6,7,8,9,10,11,12))) +
  ggtitle("Scatter Plot of Arrival Month vs Average Room Price") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data, aes(x=arrival_date, y=avg_price_per_room)) +
  geom_point(shape=1,size=1.5, colour="#D17A00", stroke = 0.5) +
  geom_smooth(method = "lm") + 
  xlab("arrival_date") + 
  ylab("avg_price_per_room") +
  ggtitle("Scatter Plot of Arrival Date vs Average Room Price") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data, aes(x=no_of_previous_cancellations, y=avg_price_per_room)) +
  geom_point(shape=1,size=1.5, colour="#D17A00", stroke = 0.4) +
  geom_smooth(method = "lm") + 
  xlab("no_of_previous_cancellations") + 
  ylab("avg_price_per_room") +
  ggtitle("Scatter Plot of Number of Previous Cancellations vs Average Room Price") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data, aes(x=no_of_previous_bookings_not_canceled, y=avg_price_per_room)) +
  geom_point(shape=1,size=1.5, colour="#D17A00", stroke = 0.5) +
  geom_smooth(method = "lm") + 
  xlab("no_of_previous_bookings_not_canceled") + 
  ylab("avg_price_per_room") +
  ggtitle("Scatter Plot of Number of not Cancelled Previous Bookings vs Average Room Price") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(data, aes(x=no_of_special_requests, y=avg_price_per_room)) +
  geom_point(shape=1,size=1.5, colour="#D17A00", stroke = 0.4) +
  geom_smooth(method = "lm") + 
  xlab("no_of_special_requests") + 
  ylab("avg_price_per_room") +
  ggtitle("Scatter Plot of Number of Special Requests vs Average Room Price") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
str(data)
```

```{r}
cor_matrix <- cor(data[sapply(data, is.numeric)], use = "complete.obs", method = "spearman")
cor_matrix

cor_melted <- melt(cor_matrix)

ggplot(cor_melted, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", name = "Correlation") +
  geom_text(aes(label = sprintf("%.2f", value)), size=3) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Numerical Variables

```{r}
numeric_variables_test_results <- data.frame(variable = character(), cor.test.coefficient = numeric(), cor.test.p.value = numeric(), lm.adj.r.sqrd = numeric(), lm.p.value = numeric(), stringsAsFactors = FALSE)
str(numeric_variables_test_results)
```

## no_of_adults

```{r}
cor.test.res <- cor.test(data$avg_price_per_room, data$no_of_adults, method="spearman")
cor.test.res
```

```{r}
lm_summary <- summary(lm(data$avg_price_per_room~data$no_of_adults))
lm_summary
```

```{r}
numeric_variables_test_results <- rbind(numeric_variables_test_results, data.frame(
  variable = "no_of_adults",
  cor.test.coefficient = cor.test.res$estimate,
  cor.test.p.value = cor.test.res$p.value,
  lm.adj.r.sqrd = lm_summary$adj.r.squared,
  lm.p.value = lm_summary$coefficients["data$no_of_adults", 4]
))
```

## no_of_children

```{r}
cor.test.res <- cor.test(data$avg_price_per_room, data$no_of_children, method="spearman")
cor.test.res
```

```{r}
lm_summary <- summary(lm(data$avg_price_per_room~data$no_of_children))
lm_summary
```

```{r}
numeric_variables_test_results <- rbind(numeric_variables_test_results, data.frame(
  variable = "no_of_children",
  cor.test.coefficient = cor.test.res$estimate,
  cor.test.p.value = cor.test.res$p.value,
  lm.adj.r.sqrd = lm_summary$adj.r.squared,
  lm.p.value = lm_summary$coefficients["data$no_of_children", 4]
))
```

## no_of_weekend_nights

```{r}
cor.test.res <- cor.test(data$avg_price_per_room, data$no_of_weekend_nights, method="spearman")
cor.test.res
```

```{r}
lm_summary <- summary(lm(data$avg_price_per_room~data$no_of_weekend_nights))
lm_summary
```

```{r}
numeric_variables_test_results <- rbind(numeric_variables_test_results, data.frame(
  variable = "no_of_weekend_nights",
  cor.test.coefficient = cor.test.res$estimate,
  cor.test.p.value = cor.test.res$p.value,
  lm.adj.r.sqrd = lm_summary$adj.r.squared,
  lm.p.value = lm_summary$coefficients["data$no_of_weekend_nights", 4]
))
```

## no_of_week_nights

```{r}
cor.test.res <- cor.test(data$avg_price_per_room, data$no_of_week_nights, method="spearman")
cor.test.res
```

```{r}
lm_summary <- summary(lm(data$avg_price_per_room~data$no_of_week_nights))
lm_summary
```

```{r}
numeric_variables_test_results <- rbind(numeric_variables_test_results, data.frame(
  variable = "no_of_week_nights",
  cor.test.coefficient = cor.test.res$estimate,
  cor.test.p.value = cor.test.res$p.value,
  lm.adj.r.sqrd = lm_summary$adj.r.squared,
  lm.p.value = lm_summary$coefficients["data$no_of_week_nights", 4]
))
```

## lead_time

```{r}
cor.test.res <- cor.test(data$avg_price_per_room, data$lead_time, method="spearman")
cor.test.res
```

```{r}
lm_summary <- summary(lm(data$avg_price_per_room~data$lead_time))
lm_summary
```

```{r}
numeric_variables_test_results <- rbind(numeric_variables_test_results, data.frame(
  variable = "lead_time",
  cor.test.coefficient = cor.test.res$estimate,
  cor.test.p.value = cor.test.res$p.value,
  lm.adj.r.sqrd = lm_summary$adj.r.squared,
  lm.p.value = lm_summary$coefficients["data$lead_time", 4]
))
```

## arrival_month

```{r}
cor.test.res <- cor.test(data$avg_price_per_room, data$arrival_month, method="spearman")
cor.test.res
```

```{r}
lm_summary <- summary(lm(data$avg_price_per_room~data$arrival_month))
lm_summary
```

```{r}
numeric_variables_test_results <- rbind(numeric_variables_test_results, data.frame(
  variable = "arrival_month",
  cor.test.coefficient = cor.test.res$estimate,
  cor.test.p.value = cor.test.res$p.value,
  lm.adj.r.sqrd = lm_summary$adj.r.squared,
  lm.p.value = lm_summary$coefficients["data$arrival_month", 4]
))
```

## arrival_date

```{r}
cor.test.res <- cor.test(data$avg_price_per_room, data$arrival_date, method="spearman")
cor.test.res
```
Above 0.05, therefore there is no evidence to reject the null hypothesis that there is no relationship between the variables.

```{r}
lm_summary <- summary(lm(data$avg_price_per_room~data$arrival_date))
lm_summary
```

```{r}
numeric_variables_test_results <- rbind(numeric_variables_test_results, data.frame(
  variable = "arrival_date",
  cor.test.coefficient = cor.test.res$estimate,
  cor.test.p.value = cor.test.res$p.value,
  lm.adj.r.sqrd = lm_summary$adj.r.squared,
  lm.p.value = lm_summary$coefficients["data$arrival_date", 4]
))
```

## no_of_previous_cancellations

```{r}
cor.test.res <- cor.test(data$avg_price_per_room, data$no_of_previous_cancellations, method="spearman")
cor.test.res
```

```{r}
lm_summary <- summary(lm(data$avg_price_per_room~data$no_of_previous_cancellations))
lm_summary
```

```{r}
numeric_variables_test_results <- rbind(numeric_variables_test_results, data.frame(
  variable = "no_of_previous_cancellations",
  cor.test.coefficient = cor.test.res$estimate,
  cor.test.p.value = cor.test.res$p.value,
  lm.adj.r.sqrd = lm_summary$adj.r.squared,
  lm.p.value = lm_summary$coefficients["data$no_of_previous_cancellations", 4]
))
```

## no_of_previous_bookings_not_canceled

```{r}
cor.test.res <- cor.test(data$avg_price_per_room, data$no_of_previous_bookings_not_canceled, method="spearman")
cor.test.res
```

```{r}
lm_summary <- summary(lm(data$avg_price_per_room~data$no_of_previous_bookings_not_canceled))
lm_summary
```

```{r}
numeric_variables_test_results <- rbind(numeric_variables_test_results, data.frame(
  variable = "no_of_previous_bookings_not_canceled",
  cor.test.coefficient = cor.test.res$estimate,
  cor.test.p.value = cor.test.res$p.value,
  lm.adj.r.sqrd = lm_summary$adj.r.squared,
  lm.p.value = lm_summary$coefficients["data$no_of_previous_bookings_not_canceled", 4]
))
```

## no_of_special_requests

```{r}
cor.test.res <- cor.test(data$avg_price_per_room, data$no_of_special_requests, method="spearman")
cor.test.res
```

```{r}
lm_summary <- summary(lm(data$avg_price_per_room~data$no_of_special_requests))
lm_summary
```

```{r}
numeric_variables_test_results <- rbind(numeric_variables_test_results, data.frame(
  variable = "no_of_special_requests",
  cor.test.coefficient = cor.test.res$estimate,
  cor.test.p.value = cor.test.res$p.value,
  lm.adj.r.sqrd = lm_summary$adj.r.squared,
  lm.p.value = lm_summary$coefficients["data$no_of_special_requests", 4]
))
```

## Evaluating Numeric Variables

```{r}
numeric_variables_test_results$cor.test.coefficient <- as.numeric(numeric_variables_test_results$cor.test.coefficient)
numeric_variables_test_results$cor.test.p.value <- as.numeric(numeric_variables_test_results$cor.test.p.value)
numeric_variables_test_results$lm.adj.r.sqrd <- as.numeric(numeric_variables_test_results$lm.adj.r.sqrd)
numeric_variables_test_results$lm.p.value <- as.numeric(numeric_variables_test_results$lm.p.value)

numeric_variables_test_results$cor.test.coefficient <- sprintf("%.4f", numeric_variables_test_results$cor.test.coefficient)
numeric_variables_test_results$cor.test.p.value <- sprintf("%.4f", numeric_variables_test_results$cor.test.p.value)
numeric_variables_test_results$lm.adj.r.sqrd <- sprintf("%.4f", numeric_variables_test_results$lm.adj.r.sqrd)
numeric_variables_test_results$lm.p.value <- sprintf("%.4f", numeric_variables_test_results$lm.p.value)

numeric_variables_test_results
```

## Categorical Variables

```{r}
names(data[sapply(data, function(x) !is.numeric(x))])
```

```{r}
categorical_variables_test_results <- data.frame(variable = character(), aov.adj.r.sqrd = numeric(), aov.p.value = numeric(), stringsAsFactors = FALSE)
```

## type_of_meal_plan

```{r}
kruskal_result <- kruskal.test(avg_price_per_room ~ type_of_meal_plan, data = data)

# Print the results
print(kruskal_result)
```


```{r}
type_of_meal_plan_anova_model <- aov(data$avg_price_per_room~data$type_of_meal_plan)
summary <- summary(type_of_meal_plan_anova_model)
summary_lm <- summary.lm(type_of_meal_plan_anova_model)
summary_lm
```


```{r}
categorical_variables_test_results <- rbind(categorical_variables_test_results, data.frame(
  variable = "type_of_meal_plan",
  aov.adj.r.sqrd = summary_lm$adj.r.squared,
  aov.p.value = summary[[1]][["Pr(>F)"]][1]
))
```

## required_car_parking_space

```{r}
kruskal_result <- kruskal.test(avg_price_per_room ~ required_car_parking_space, data = data)

# Print the results
print(kruskal_result)
```

```{r}
required_car_parking_space_anova_model <- aov(data$avg_price_per_room~data$required_car_parking_space)
summary <- summary(required_car_parking_space_anova_model)
summary_lm <- summary.lm(required_car_parking_space_anova_model)
summary_lm
```


```{r}
categorical_variables_test_results <- rbind(categorical_variables_test_results, data.frame(
  variable = "required_car_parking_space",
  aov.adj.r.sqrd = summary_lm$adj.r.squared,
  aov.p.value = summary[[1]][["Pr(>F)"]][1]
))
```

## room_type_reserved

```{r}
kruskal_result <- kruskal.test(avg_price_per_room ~ room_type_reserved, data = data)

# Print the results
print(kruskal_result)
```


```{r}
room_type_reserved_anova_model <- aov(data$avg_price_per_room~data$room_type_reserved)
summary <- summary(room_type_reserved_anova_model)
summary_lm <- summary.lm(room_type_reserved_anova_model)
summary_lm
```

```{r}
categorical_variables_test_results <- rbind(categorical_variables_test_results, data.frame(
  variable = "room_type_reserved",
  aov.adj.r.sqrd = summary_lm$adj.r.squared,
  aov.p.value = summary[[1]][["Pr(>F)"]][1]
))
```

## market_segment_type

```{r}
kruskal_result <- kruskal.test(avg_price_per_room ~ market_segment_type, data = data)

# Print the results
print(kruskal_result)
```


```{r}
market_segment_type_anova_model <- aov(data$avg_price_per_room~data$market_segment_type)
summary <- summary(market_segment_type_anova_model)
summary_lm <- summary.lm(market_segment_type_anova_model)
summary_lm
```

```{r}
categorical_variables_test_results <- rbind(categorical_variables_test_results, data.frame(
  variable = "market_segment_type",
  aov.adj.r.sqrd = summary_lm$adj.r.squared,
  aov.p.value = summary[[1]][["Pr(>F)"]][1]
))
```

## repeated_guest

```{r}
repeated_guest_anova_model <- aov(data$avg_price_per_room~data$repeated_guest)
summary <- summary(repeated_guest_anova_model)
summary_lm <- summary.lm(repeated_guest_anova_model)
summary_lm
```

```{r}
categorical_variables_test_results <- rbind(categorical_variables_test_results, data.frame(
  variable = "repeated_guest",
  aov.adj.r.sqrd = summary_lm$adj.r.squared,
  aov.p.value = summary[[1]][["Pr(>F)"]][1]
))
```

## booking_status

```{r}
booking_status_anova_model <- aov(data$avg_price_per_room~data$booking_status)
summary <- summary(booking_status_anova_model)
summary_lm <- summary.lm(booking_status_anova_model)
summary_lm
```

```{r}
categorical_variables_test_results <- rbind(categorical_variables_test_results, data.frame(
  variable = "booking_status",
  aov.adj.r.sqrd = summary_lm$adj.r.squared,
  aov.p.value = summary[[1]][["Pr(>F)"]][1]
))
```

## arrival_year

```{r}
arrival_year_anova_model <- aov(data$avg_price_per_room~data$arrival_year)
summary <- summary(arrival_year_anova_model)
summary_lm <- summary.lm(arrival_year_anova_model)
summary_lm
```
        
        
```{r}
wilcox.test.res <- wilcox.test(data$avg_price_per_room~data$arrival_year)
wilcox.test.res
```

        
        
```{r}
categorical_variables_test_results <- rbind(categorical_variables_test_results, data.frame(
  variable = "arrival_year",
  aov.adj.r.sqrd = summary_lm$adj.r.squared,
  aov.p.value = summary[[1]][["Pr(>F)"]][1]
))
```


## Evaluating Categorical Variables

```{r}
categorical_variables_test_results$aov.adj.r.sqrd <- as.numeric(categorical_variables_test_results$aov.adj.r.sqrd)
categorical_variables_test_results$aov.p.value <- as.numeric(categorical_variables_test_results$aov.p.value)

categorical_variables_test_results$aov.adj.r.sqrd <- sprintf("%.4f", categorical_variables_test_results$aov.adj.r.sqrd)
categorical_variables_test_results$aov.p.value <- sprintf("%.4f", categorical_variables_test_results$aov.p.value)

categorical_variables_test_results
```

## Logical Variables

```{r}
names(data[sapply(data, is.logical)])
```

```{r}
logical_variables_test_results <- data.frame(variable = character(), log.model.aic = numeric(), t.test.p.value = numeric(), wolcox.test.p.value = numeric(), stringsAsFactors = FALSE)
```

## required_car_parking_space

```{r}
required_car_parking_space_log_reg_model <- glm(data$required_car_parking_space~data$avg_price_per_room, family = "binomial")
summary(required_car_parking_space_log_reg_model)
```

```{r}
t.test.res <- t.test(data$avg_price_per_room~data$required_car_parking_space)
t.test.res
```

```{r}
wilcox.test.res <- wilcox.test(data$avg_price_per_room~data$required_car_parking_space)
wilcox.test.res
```

```{r}
logical_variables_test_results <- rbind(logical_variables_test_results, data.frame(
  variable = "required_car_parking_space",
  log.model.aic = AIC(required_car_parking_space_log_reg_model),
  t.test.p.value = t.test.res$p.value,
  wolcox.test.p.value = wilcox.test.res$p.value
))
```

## repeated_guest

```{r}
repeated_guest_log_reg_model <- glm(data$repeated_guest~data$avg_price_per_room, family = "binomial")
summary(repeated_guest_log_reg_model)
```

```{r}
t.test.res <- t.test(data$avg_price_per_room~data$repeated_guest)
t.test.res
```

```{r}
wilcox.test.res <- wilcox.test(data$avg_price_per_room~data$repeated_guest)
wilcox.test.res
```

```{r}
logical_variables_test_results <- rbind(logical_variables_test_results, data.frame(
  variable = "repeated_guest",
  log.model.aic = AIC(repeated_guest_log_reg_model),
  t.test.p.value = t.test.res$p.value,
  wolcox.test.p.value = wilcox.test.res$p.value
))
```

## booking_status

```{r}
booking_status_log_reg_model <- glm(data$booking_status~data$avg_price_per_room, family = "binomial")
summary(booking_status_log_reg_model)
```

```{r}
t.test.res <- t.test(data$avg_price_per_room~data$booking_status)
t.test.res
```

```{r}
wilcox.test.res <- wilcox.test(data$avg_price_per_room~data$booking_status)
wilcox.test.res
```


```{r}
logical_variables_test_results <- rbind(logical_variables_test_results, data.frame(
  variable = "booking_status",
  log.model.aic = AIC(booking_status_log_reg_model),
  t.test.p.value = t.test.res$p.value,
  wolcox.test.p.value = wilcox.test.res$p.value
))
```

## Evaluating Logical Variables

```{r}
logical_variables_test_results$log.model.aic <- as.numeric(logical_variables_test_results$log.model.aic)
logical_variables_test_results$t.test.p.value <- as.numeric(logical_variables_test_results$t.test.p.value)
logical_variables_test_results$wolcox.test.p.value <- as.numeric(logical_variables_test_results$wolcox.test.p.value)

logical_variables_test_results$log.model.aic <- sprintf("%.4f", logical_variables_test_results$log.model.aic)
logical_variables_test_results$t.test.p.value <- sprintf("%.4f", logical_variables_test_results$t.test.p.value)
logical_variables_test_results$wolcox.test.p.value <- sprintf("%.4f", logical_variables_test_results$wolcox.test.p.value)

logical_variables_test_results
```

```{r}
# Remove Booking_ID from Dataset
data <- data[ , !(names(data) %in% c("Booking_ID"))]
```

```{r}
str(data)
```

## PCA

```{r}
#Turn characters to factors
data$market_segment_type <- as.factor(data$market_segment_type)
data$room_type_reserved <- as.factor(data$room_type_reserved)

# Identify categorical variables
categorical_vars <- sapply(data, is.factor) | sapply(data, is.character)

# One-hot encode categorical variables
data_encoded <- model.matrix(~ . -1, data = data[, categorical_vars], contrasts.arg = lapply(data[sapply(data, is.factor)], contrasts, contrasts=FALSE))


# Convert to data.frame
data_encoded <- as.data.frame(data_encoded)


# Extract numeric variables
numeric_data <- data[, !categorical_vars]

# Combine numeric and encoded categorical variables
combined_data <- cbind(numeric_data, data_encoded)
data_numeric <- combined_data

# Remove Output Variable
combined_data_no_output = combined_data[, !names(combined_data) %in% "avg_price_per_room"]

# Apply PCA
pca_result <- prcomp(combined_data_no_output, center = TRUE, scale. = TRUE)

# Save PCA Scores
write.csv(pca_result$x, file = "pca_scores1.csv")

# View summary of PCA results
summary(pca_result)

pca_correlation_matrix <- cor(combined_data, pca_result$x)

pca_correlation_matrix_melted_1 = melt(pca_correlation_matrix[1:nrow(pca_correlation_matrix)/2, 1:8])
pca_correlation_matrix_melted_2 = melt(pca_correlation_matrix[((nrow(pca_correlation_matrix)/2) + 1):nrow(pca_correlation_matrix), 1:8])

ggplot(pca_correlation_matrix_melted_1, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", name = "Correlation") +
  geom_text(aes(label = sprintf("%.2f", value)), size = 3.25) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(pca_correlation_matrix_melted_2, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", name = "Correlation") +
  geom_text(aes(label = sprintf("%.2f", value)), size = 3.25) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8))
```




```{r}
pca_result_var <- pca_result$sdev^2
pca_result_PEV <- pca_result_var / sum(pca_result_var)

opar <- par(no.readonly = TRUE)
plot(
  cumsum(pca_result_PEV),
  ylim = c(0,1),
  xlab = 'PC',
  ylab = 'cumulative PEV',
  pch = 20,
  col = '#0044ff'
)
abline(h = 0.8, col = 'red', lty = 'dashed')
par(opar)

# PCA Biplot
scores <- as.data.frame(pca_result$x[, 1:2])
loadings <- as.data.frame(pca_result$rotation[, 1:2])
loadings$variable <- rownames(loadings)

pca_biplot <- ggplot() +
  geom_point(data = scores, aes(x = PC1, y = PC2), colour = '#FDBCB4', size = 1.5, shape = 1, stroke = 0.7) +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "white", colour = "white"), plot.margin = unit(c(0.5, 0.1, 0.1, 0.05), 
                                "inches")) +
  theme(plot.title = element_text(hjust = 0.5, vjust = 7, family = "Arial" )) +
  labs(title = "PCA Biplot", x = "PC1", y = "PC2") +
  geom_segment(data = loadings, aes(x = 0, y = 0, xend = PC1 * max(abs(scores$PC1)), yend = PC2 * max(abs(scores$PC2))), 
                      arrow = arrow(type = "closed", length = unit(0.1, "inches"), angle = 18), color = "#555555", size = 0.5, alpha = 0.65) +
  geom_text(data = loadings, aes(x = PC1 * max(abs(scores$PC1)), y = PC2 * max(abs(scores$PC2)), label = variable), 
            position = position_nudge(x = 0.1, y = 0.1), check_overlap = TRUE, hjust = 0, vjust = 0, color = "#1B4B50", size = 3.27) 

ggsave("PC1_and_PC2_Plot1.png", plot = pca_biplot, width = 10, height = 10, dpi = 300, bg="white")
dev.off()
print(pca_biplot)
```


## Clustering

```{r}
k_data = kmeans(combined_data_no_output, 12)

k_cluster_id_data <- k_data$cluster

data$cluster <- k_cluster_id_data
data_numeric$cluster <- k_cluster_id_data
```

```{r}
cluster_anova_model <- aov(data$avg_price_per_room ~ factor(data$cluster))
summary <- summary(cluster_anova_model)
summary_lm <- summary.lm(cluster_anova_model)
summary_lm
```

```{r}
categorical_variables_test_results <- rbind(categorical_variables_test_results, data.frame(
  variable = "cluster",
  aov.adj.r.sqrd = summary_lm$adj.r.squared,
  aov.p.value = summary[[1]][["Pr(>F)"]][1]
))

categorical_variables_test_results$aov.adj.r.sqrd <- as.numeric(categorical_variables_test_results$aov.adj.r.sqrd)
categorical_variables_test_results$aov.p.value <- as.numeric(categorical_variables_test_results$aov.p.value)

categorical_variables_test_results$aov.adj.r.sqrd <- sprintf("%.4f", categorical_variables_test_results$aov.adj.r.sqrd)
categorical_variables_test_results$aov.p.value <- sprintf("%.4f", categorical_variables_test_results$aov.p.value)
```

## Saving Data

```{r}
logical_cols <- sapply(data_numeric, is.logical)
data_numeric[logical_cols] <- lapply(data_numeric[logical_cols], as.integer)
```

```{r}
write.csv(data, file = "post_eda_dataset1.csv", row.names = FALSE)
```

```{r}
write.csv(data_numeric, file = "post_eda_numerical_dataset1.csv", row.names = FALSE)
```

## Comparing Original Variables to PCA Variables

```{r}
original_data <- read_csv("post_eda_dataset1.csv")
pca_data <- read_csv("pca_scores1.csv")
```

```{r}
# For the original dataset
y <- original_data$avg_price_per_room
X_original <- original_data[, !names(original_data) %in% "avg_price_per_room"]
# X_original <- original_data[, -ncol(original_data)]  # Exclude the last column

# For the PCA dataset (assuming no target variable included)
X_pca <- pca_data
```

```{r}
set.seed(42)  # For reproducibility
index <- createDataPartition(y, p = .8, list = FALSE)
X_train_orig <- X_original[index, ]
X_test_orig <- X_original[-index, ]
X_train_pca <- X_pca[index, ]
X_test_pca <- X_pca[-index, ]
y_train <- y[index]
y_test <- y[-index]
```

```{r}
# Decision Tree on Original Data
model_dt_orig <- rpart(y_train ~ ., data = as.data.frame(X_train_orig))
predictions_dt_orig <- predict(model_dt_orig, newdata = as.data.frame(X_test_orig))
rmse_dt_orig <- rmse(predictions_dt_orig, y_test)

# Decision Tree on PCA Data
model_dt_pca <- rpart(y_train ~ ., data = as.data.frame(X_train_pca))
predictions_dt_pca <- predict(model_dt_pca, newdata = as.data.frame(X_test_pca))
rmse_dt_pca <- rmse(predictions_dt_pca, y_test)

# Output RMSE values
print(paste("Decision Tree RMSE on Original Data:", rmse_dt_orig))
print(paste("Decision Tree RMSE on PCA Data:", rmse_dt_pca))
```

```{r}
# SVM on Original Data
model_svm_orig <- svm(y_train ~ ., data = data.frame(X_train_orig, y_train))
predictions_svm_orig <- predict(model_svm_orig, newdata = X_test_orig)
rmse_svm_orig <- rmse(y_test, predictions_svm_orig)

# SVM on PCA Data
model_svm_pca <- svm(y_train ~ ., data = data.frame(X_train_pca, y_train))
predictions_svm_pca <- predict(model_svm_pca, newdata = X_test_pca)
rmse_svm_pca <- rmse(y_test, predictions_svm_pca)

# Print RMSE for SVMs
print(paste("SVM RMSE on Original Data:", rmse_svm_orig))
print(paste("SVM RMSE on PCA Data:", rmse_svm_pca))
```

```{r}
# Assuming X_train_orig, X_test_orig, y_train, and y_test are already defined
rf_model_orig <- randomForest(x = as.data.frame(X_train_orig), y = y_train)
predictions_rf_orig <- predict(rf_model_orig, newdata = as.data.frame(X_test_orig))
rmse_rf_orig <- sqrt(mean((predictions_rf_orig - y_test)^2))

# Assuming X_train_pca, X_test_pca, y_train, and y_test are already defined
rf_model_pca <- randomForest(x = as.data.frame(X_train_pca), y = y_train)
predictions_rf_pca <- predict(rf_model_pca, newdata = as.data.frame(X_test_pca))
rmse_rf_pca <- sqrt(mean((predictions_rf_pca - y_test)^2))

print(paste("Random Forest RMSE on Original Data:", rmse_rf_orig))
print(paste("Random Forest RMSE on PCA Data:", rmse_rf_pca))
```

## Getting Loss of Random Forest Models given a Single Var + PC1 + PC2

```{r}
set.seed(64)

getRandomForestLossForSingleVar <- function(X, y, var_name) {
  if (var_name == "None") {
    X <- select(pca_data, "PC1", "PC2")
  } else {
    X <- bind_cols(
      X %>% select(all_of(var_name)),
      select(pca_data, "PC1", "PC2")
    )
  }
  data_split_index <- createDataPartition(y, p = .8, list = FALSE)
  
  X_train <- data.frame(X = X[data_split_index,])
  X_test <- data.frame(X = X[-data_split_index,])
  y_train <- y[data_split_index]
  y_test <- y[-data_split_index]
  
  model <- randomForest(x = X_train, y = y_train)
  
  predictions <- predict(model, newdata = X_test)
  test_rmse <- sqrt(mean((predictions - y_test)^2))
  return(test_rmse)
}

data_no_target <- data[, !names(data) %in% "avg_price_per_room"]
random_forest_loss_for_variables <- data.frame(Variable = character(), Loss = numeric(), stringsAsFactors = FALSE)

test_rmse <- getRandomForestLossForSingleVar(data, data$avg_price_per_room, "None")
random_forest_loss_for_variables <- rbind(random_forest_loss_for_variables, data.frame(Variable = "None", Loss = test_rmse))

for(column_name in colnames(data_no_target)) {
  test_rmse <- getRandomForestLossForSingleVar(data, data$avg_price_per_room, column_name)
  random_forest_loss_for_variables <- rbind(random_forest_loss_for_variables, data.frame(Variable = column_name, Loss = test_rmse))
  cat("Built Model for ", grep(column_name, colnames(data_no_target)), " out of ", length(colnames(data_no_target)), "Variables\n")
}

random_forest_loss_for_variables <- random_forest_loss_for_variables %>% arrange(Loss)
random_forest_loss_for_variables
```

## Reviewing Variables Test Results

```{r}
numeric_variables_test_results
categorical_variables_test_results
logical_variables_test_results
```

